---
title: "Trends and Topics of Statistics Training Courses"
author: "Mark Andrews, Eirini Koutoumanou, Matt Castle"
format: revealjs
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning = FALSE, message = FALSE)
```

## Introduction

* The aim of this work is to analyse what is being taught, to whom, how, etc. in statistics training courses (usually CPD, advanced training for research).
* Why this might be of general relevance for statistics education:
  - Training courses reveal, and help close, teaching-workplace skills/knowledge gap
  - They are reflect evolving or emerging methods and tools
  - They are a principal means of teacher-training/upskilling

## Databases

```{r}
library(here)
library(tidyverse)

allstat_n <- readRDS(here('data/allstat_training_course_posts_ollama.Rds'))$data_frame |> nrow()
tess_n <- readRDS(here('data/tess_training_course_posts_llama_results.Rds'))$data_frame |> nrow()
ncrm_n <- readRDS(here('data/ncrm_training_course_posts_llama_results.Rds'))$data_frame |> nrow()

source(here('analysis/results_topics.R'))
source(here('analysis/results_fields.R'))
source(here('analysis/results_software.R'))
source(here('analysis/results_online.R'))
source(here('analysis/results_level.R'))
source(here('analysis/results_provider.R'))
source(here('analysis/results_duration.R'))
```

In total, we analysed `r format(allstat_n + tess_n + ncrm_n, big.mark = ",", scientific = FALSE)` training courses descriptions:

- ALLSTAT mailing list: `r allstat_n` posts from 1998 to 2025
- NCRM training courses and events database: `r ncrm_n` entries from 2003 to 2025
- Elixir TeSS training course database: `r tess_n` entries from 2011 to 2025

All of these descriptions were webscraped from the three websites using Python.

## Extracted information

For each training course description, we extracted

- One line description of course
- One term summary description of topic
- Topic keywords
- Intended audience (e.g. academic/research field) and level
- Software
- Duration, delivery method
- Course provider

## Extraction method

* Information was extracted using a locally running large language model (LLM), Llama 3.3.
* The LLM ran on a Linux workstation with a RTX A6000 GPU, 10,752 cuda cores and 48GB of VRAM.
* The script was written in R using the `ellmer` package.
* Approximately 2500 descriptions processed per day.
* LLM API services provided by OpenAI could easily have been used (one line change in R code), approximately £50.

## Topics

```{r}
topics_df |> top_categories() |> 
  mutate(proportion = n/sum(n)) |> 
  slice_head(n=10) |> 
  ggplot(aes(x = reorder(value, n), y = proportion)) + geom_col() +
  coord_flip() + xlab(NULL) +
  theme_classic()
```


## Topics (ALLSTAT)

```{r}
topics_df |>
  filter(db == 'ALLSTAT') |> 
  top_categories() |> 
  mutate(proportion = n/sum(n)) |> 
  slice_head(n=10) |> 
  ggplot(aes(x = reorder(value, n), y = proportion)) + geom_col() +
  coord_flip() + xlab(NULL) +
  theme_classic()
```


## Topics (NCRM)

```{r}
topics_df |>
  filter(db == 'NCRM') |> 
  top_categories() |> 
  mutate(proportion = n/sum(n)) |> 
  slice_head(n=10) |> 
  ggplot(aes(x = reorder(value, n), y = proportion)) + geom_col() +
  coord_flip() + xlab(NULL) +
  theme_classic()
```

## Topics (most recent quintile)

```{r}
topics_df |>
  filter(year_group == 5) |> 
  top_categories() |> 
  mutate(proportion = n/sum(n)) |> 
  slice_head(n=10) |> 
  ggplot(aes(x = reorder(value, n), y = proportion)) + geom_col() +
  coord_flip() + xlab(NULL) +
  theme_classic()
```

## Topics (least recent quintile)

```{r}
topics_df |>
  filter(year_group == 1) |> 
  top_categories() |> 
  mutate(proportion = n/sum(n)) |> 
  slice_head(n=10) |> 
  ggplot(aes(x = reorder(value, n), y = proportion)) + geom_col() +
  coord_flip() + xlab(NULL) +
  theme_classic()
```

## Intended academic/research field

```{r}
fields_df |>
  count(field) |> 
  arrange(desc(n)) |> 
  mutate(proportion = n/sum(n)) |> 
  slice_head(n=10) |> 
  ggplot(aes(x = reorder(field, n), y = proportion)) + geom_col() +
  coord_flip() + xlab(NULL) +
  theme_classic()
```

## Intended academic/research field (ALLSTAT)

```{r}
fields_df |>
  filter(db == 'ALLSTAT') |> 
  count(field) |> 
  arrange(desc(n)) |> 
  mutate(proportion = n/sum(n)) |> 
  slice_head(n=10) |> 
  ggplot(aes(x = reorder(field, n), y = proportion)) + geom_col() +
  coord_flip() + xlab(NULL) +
  theme_classic()
```

## Intended academic/research field (NCRM)

```{r}
fields_df |>
  filter(db == 'NCRM') |> 
  count(field) |> 
  arrange(desc(n)) |> 
  mutate(proportion = n/sum(n)) |> 
  slice_head(n=10) |> 
  ggplot(aes(x = reorder(field, n), y = proportion)) + geom_col() +
  coord_flip() + xlab(NULL) +
  theme_classic()
```

## Intended academic/research field (TESS)

```{r}
fields_df |>
  filter(db == 'TESS') |> 
  count(field) |> 
  arrange(desc(n)) |> 
  mutate(proportion = n/sum(n)) |> 
  slice_head(n=10) |> 
  ggplot(aes(x = reorder(field, n), y = proportion)) + geom_col() +
  coord_flip() + xlab(NULL) +
  theme_classic()
```

## Intended level

```{r}
level_df |>
  count(level) |> 
  arrange(desc(n)) |> 
  mutate(proportion = n/sum(n)) |> 
  slice_head(n=10) |> 
  ggplot(aes(x = reorder(level, n), y = proportion)) + geom_col() +
  coord_flip() + xlab(NULL) +
  theme_classic()
```

## Software

```{r}
software_df |>
  count(software) |> 
  arrange(desc(n)) |> 
  mutate(proportion = n/sum(n)) |> 
  slice_head(n=10) |> 
  ggplot(aes(x = reorder(software, n), y = proportion)) + geom_col() +
  coord_flip() + xlab(NULL) +
  theme_classic()
```

## Software (year quintile 1, least recent)

```{r}
software_df |>
  filter(year_group == 1) |> 
  count(software) |> 
  arrange(desc(n)) |> 
  mutate(proportion = n/sum(n)) |> 
  slice_head(n=10) |> 
  ggplot(aes(x = reorder(software, n), y = proportion)) + geom_col() +
  coord_flip() + xlab(NULL) +
  theme_classic()
```

<!-- ## Software (year quintile 2) -->

<!-- ```{r} -->
<!-- software_df |> -->
<!--   filter(year_group == 2) |>  -->
<!--   count(software) |>  -->
<!--   arrange(desc(n)) |>  -->
<!--   mutate(proportion = n/sum(n)) |>  -->
<!--   slice_head(n=10) |>  -->
<!--   ggplot(aes(x = reorder(software, n), y = proportion)) + geom_col() + -->
<!--   coord_flip() + xlab(NULL) + -->
<!--   theme_classic() -->
<!-- ``` -->

<!-- ## Software (year quintile 3) -->

<!-- ```{r} -->
<!-- software_df |> -->
<!--   filter(year_group == 3) |>  -->
<!--   count(software) |>  -->
<!--   arrange(desc(n)) |>  -->
<!--   mutate(proportion = n/sum(n)) |>  -->
<!--   slice_head(n=10) |>  -->
<!--   ggplot(aes(x = reorder(software, n), y = proportion)) + geom_col() + -->
<!--   coord_flip() + xlab(NULL) + -->
<!--   theme_classic() -->
<!-- ``` -->

<!-- ## Software (year quintile 4) -->

<!-- ```{r} -->
<!-- software_df |> -->
<!--   filter(year_group == 4) |>  -->
<!--   count(software) |>  -->
<!--   arrange(desc(n)) |>  -->
<!--   mutate(proportion = n/sum(n)) |>  -->
<!--   slice_head(n=10) |>  -->
<!--   ggplot(aes(x = reorder(software, n), y = proportion)) + geom_col() + -->
<!--   coord_flip() + xlab(NULL) + -->
<!--   theme_classic() -->
<!-- ``` -->

## Software (year quintile 5, most recent)

```{r}
software_df |>
  filter(year_group == 5) |> 
  count(software) |> 
  arrange(desc(n)) |> 
  mutate(proportion = n/sum(n)) |> 
  slice_head(n=10) |> 
  ggplot(aes(x = reorder(software, n), y = proportion)) + geom_col() +
  coord_flip() + xlab(NULL) +
  theme_classic()
```

## Delivery method

```{r}
delivery_df |>
  filter(delivery == 'online' | delivery == 'in person') |> 
  # filter(year_group == 5) |> 
  count(delivery) |> 
  arrange(desc(n)) |> 
  mutate(proportion = n/sum(n)) |> 
  slice_head(n=10) |> 
  ggplot(aes(x = reorder(delivery, n), y = proportion)) + geom_col() +
  coord_flip() + xlab(NULL) +
  theme_classic()
```

## Delivery method (most recent quintile)

```{r}
delivery_df |>
  filter(delivery == 'online' | delivery == 'in person') |> 
  filter(year_group == 5) |> 
  count(delivery) |> 
  arrange(desc(n)) |> 
  mutate(proportion = n/sum(n)) |> 
  slice_head(n=10) |> 
  ggplot(aes(x = reorder(delivery, n), y = proportion)) + geom_col() +
  coord_flip() + xlab(NULL) +
  theme_classic()
```

## Delivery method (all but most recent quintile)

```{r}
delivery_df |>
  filter(delivery == 'online' | delivery == 'in person') |> 
  filter(year_group != 5) |> 
  count(delivery) |> 
  arrange(desc(n)) |> 
  mutate(proportion = n/sum(n)) |> 
  slice_head(n=10) |> 
  ggplot(aes(x = reorder(delivery, n), y = proportion)) + geom_col() +
  coord_flip() + xlab(NULL) +
  theme_classic()
```

## Duration

```{r}
duration_df |>
  count(duration) |> 
  arrange(desc(n)) |> 
  mutate(proportion = n/sum(n)) |> 
  slice_head(n=10) |> 
  ggplot(aes(x = reorder(duration, n), y = proportion)) + geom_col() +
  coord_flip() + xlab(NULL) +
  theme_classic()
```

## Provider

```{r}
provider_df |>
  count(provider) |> 
  arrange(desc(n)) |> 
  mutate(proportion = n/sum(n)) |> 
  filter(provider != 'not specified', provider != 'unknown') |> 
  slice_head(n=10) |> 
  ggplot(aes(x = reorder(provider, n), y = proportion)) + geom_col() +
  coord_flip() + xlab(NULL) +
  theme_classic()
```

